#  Reproducible Research
## Assignment 1

Read the data into a data table.

```{r, echo=TRUE}
d <- read.table('repdata_data_activity/activity.csv', sep = ',', header = TRUE, na.strings = 'NA')
origNames <- names(d)
d$datetime = strptime(d$date, '%Y-%m-%d')
```

Calculate the total steps taken in a day.

```{r, echo=TRUE}
tot <- aggregate(steps ~ date, d, sum)
names(tot) <- c("Date", "Steps")
head(tot)
```


Create a histogram

```{r, echo=TRUE}
tot$datetime = strptime(tot$Date, '%Y-%m-%d')
barplot(tot$Steps, space = 0)
```


Mean and median

```{r, echo=TRUE}
stepsmean <- mean(tot$Steps)
print(stepsmean)
stepsmedian <- median(tot$Steps)
print(stepsmedian)
```

Plot the average number of steps in each interval across all days

```{r, echo=TRUE}
avg <- aggregate(d$steps ~ d$interval, d, mean)
names(avg) <- c("Interval", "Mean")
plot(avg$Interval, avg$Mean, type = "l", xlab = "Time", ylab = "Average Steps")
```

Maximum number of steps in a 5 minute interval
```{r, echo=TRUE}
print(avg[avg$Mean == max(avg$Mean),])
```


Number of rows with missing data

```{r, echo=TRUE}
NAs <- d[is.na(d$steps),]
print(nrow(NAs))
```

Impute missing data using median calculated from available data

```{r, echo=TRUE}
med <- aggregate(d$steps ~ d$interval, d, median)
names(med) <- c("Interval", "Median")
dWithMed <-merge(d, med, by.x = "interval", by.y ="Interval")
dWithMed$imputedSteps = ifelse(is.na(dWithMed$steps), dWithMed$Median, dWithMed$steps)
ordereddata <- dWithMed[order(dWithMed$date, dWithMed$interval), ]
Imputeddata <- ordereddata[, c(6, 3, 1)]
names(Imputeddata) <- origNames
head(Imputeddata)
```

Create a histogram

```{r, echo=TRUE}
tot2 <- aggregate(steps ~ date, Imputeddata, sum)
names(tot2) <- c("Date", "Steps")
barplot(tot2$Steps, space = 0)
```

Mean and median

```{r, echo=TRUE}
print(mean(tot2$Steps))
print(median(tot2$Steps))
```

Compare activity during weekdays vs. weekends

```{r, echo=TRUE}
Imputeddata$datetime = strptime(Imputeddata$date, '%Y-%m-%d')
Imputeddata$day <- weekdays(Imputeddata$datetime)
Imputeddata$weekend <- ifelse(Imputeddata$day=="Saturday" | Imputeddata$day=="Sunday","Y","N")

avgWeekDay <- aggregate(steps ~ interval, Imputeddata[Imputeddata$weekend == 'N',], mean)
avgWeekDay$weekend = as.factor('Weekday')
avgWeekEnd <- aggregate(steps ~ interval, Imputeddata[Imputeddata$weekend == 'Y',], mean)
avgWeekEnd$weekend = as.factor('Weekend')
avg2 = rbind(avgWeekDay,avgWeekEnd)
head(avg2)

library(lattice)
xyplot(steps~interval|weekend,
       data=avg2,
       type="l",
       layout=c(1,2),
       xlab = "Time",
       ylab = "Average Steps",
)
```
